Generalization Errors of Hard Parameter Sharing: High-dimensional Asymptotics and Implications

### 2020-09-13 arxiv/proof_main.tex
sec b.2

\medskip
\noindent{\bf Deriving the variance asymptotics.} Using definition \eqref{mainG}, we can write equation \eqref{eigen2extra} as
\be\label{rewrite X as R} [(X^{(1)})^\top X^{(1)}+(X^{(2)})^\top X^{(2)}]^{-1}=n^{-1}\Sigma_2^{-1/2}V\cal G(0)V^\top\Sigma_2^{-1/2}.\ee
When $z=0$, it is easy to check that equation \eqref{selfomega_a} can be reduced to \eqref{eq_a12extra}, which means that we actually have $a_1(0)=a_1$ and $a_2(0)=a_2$. Hence the matrix limit of $\cal G(0)$ is given by $(a_{1}\Lambda^2 + a_{2}\id_p)^{-1}$. Then inserting this limit into \eqref{rewrite X as R}, we can write the left-hand side of \eqref{lem_cov_shift_eq} as
\begin{align*}
&\bigtr{(X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} \Sigma}\approx n^{-1}\bigtr{\Sigma_2^{-1/2}V\cal (a_{1}\Lambda^2 + a_{2}\id_p)^{-1}V^\top\Sigma_2^{-1/2}\Sigma}  \\
&=n^{-1}\bigtr{\Sigma_2^{-1/2}\cal (a_{1}\Sigma_2^{-1/2}\Sigma_1\Sigma_2^{-1/2} + a_{2}\id_p)^{-1}\Sigma_2^{-1/2}\Sigma}  = n^{-1}\bigtr{\cal (a_{1} \Sigma_1  + a_{2}\Sigma_2)^{-1}\Sigma}  ,
\end{align*}
where in the second step we used $V \Lambda^2 V^\top=M^\top M=\Sigma_2^{-1/2}\Sigma_1\Sigma_2^{-1/2}$ by equation \eqref{eigen2}.  This concludes \eqref{lem_cov_shift_eq} up to a small error.

@@

\medskip
\noindent{\bf Deriving the bias asymptotics.} Finally, we describe how to derive the bias asymptotics. We can write the left-hand side of equation \eqref{lem_cov_derv_eq} using the derivative of $\cal G$ with respect to $z$ at $z=0$. More precisely, using equation \eqref{eigen2extra} we can obtain that
\begin{align}
&n^2\bignorm{\Sigma_2^{1/2} \bigbrace{ (X^{(1)})^{\top}X^{(1)} + (X^{(2)})^{\top}X^{(2)} }^{-1} \Sigma_1^{1/2} w}^2 \nonumber\\
&=n^2 w^\top \Sigma_1^{1/2}\Sigma_2^{-1/2}V\left(   \Lambda U^\top (Z^{(1)})^\top Z^{(1)} U\Lambda  + V^\top (Z^{(2)})^\top Z^{(2)}V\right)^{-2}V^\top\Sigma_2^{-1/2}\Sigma_1^{1/2}w \nonumber\\
&=  w^\top \Sigma_1^{1/2}\Sigma_2^{-1/2}V\cal G'(0)V^\top\Sigma_2^{-1/2}\Sigma_1^{1/2}w,\label{calculate G'}
\end{align}
where we used equation \eqref{rewrite X as R} in the second step. Since the matrix limit of $\cal G(z)$ is given by \eqref{matrix limit}, it is natural to guess that the matrix limit of $\cal G'(0)$ is given by
\be\label{cal G'0}\cal G'(0) \approx \left.\frac{\dd}{\dd z}\right|_{z=0}(-z\id_p+a_{1}(z)\Lambda^2 + a_{2}(z)\id_p)^{-1} = \frac{\id_p- a_1'(0)\Lambda^2 - a_2'(0)\id_p}{(a_{1}(0)\Lambda^2 + a_{2}(0)\id_p)^2}.\ee
If we let $a_3:=-a_1'(0)$ and $a_4:=-a_2'(0)$, then taking implicit differentiation of equation \eqref{selfomega_a} we can check that $(a_3,a_4)$ satisfies equation \eqref{eq_a34extra}. Then inserting \eqref{cal G'0} into \eqref{calculate G'}, we obtain that
\begin{align*}
& n^2\bignorm{\Sigma_2^{1/2} \bigbrace{ (X^{(1)})^{\top}(X^{(1)}) + (X^{(2)})^{\top}(X^{(2)}) }^{-1} \Sigma_1^{1/2} w}^2\\
&\approx  w^\top \Sigma_1^{1/2}\Sigma_2^{-1/2}V\frac{a_3\Lambda^2 +(1+ a_4)\id_p}{(a_{1}\Lambda^2 + a_{2}\id_p)^2}V^\top \Sigma_2^{-1/2}\Sigma_1^{1/2}w= w^{\top} \Pi w,
\end{align*}
where in the last step we used $M = \Sigma_1^{1/2}\Sigma_2^{-1/2}$ and $V \Lambda^2 V^\top=M^\top M$. This concludes \eqref{lem_cov_derv_eq}. Note that in order to have the approximate identity for $\cal G'(0)$ in \eqref{cal G'0}, we not only need to know the asymptotics of $\cal G(0)$, but also need to know the asymptotics of $\cal G(z)$ for general $z$ around $z=0$. This is the main reason why we need to take a general $z$ in the definition of resolvents.
%In the above definition, we have taken the argument of $\cal G$ to be a general complex number, because we will need to use $\cal G'(0)$ in the proof of Lemma \ref{lem_cov_derivative}, which requires a good estimate of $\cal G(z)$ for $z$ around the origin.
%For the variance asymptotic limit, we study the resolvent
%	\[ R(z):= \bigbrace{\Sigma_2^{-1/2}( (X^{(1)})^{\top}(X^{(1)}) + (X^{(2)})^{\top}(X^{(2)}))\Sigma_2^{-1/2} - z \id}^{-1}, \text{ for any } z\in \C \text{ around } z=0. \]
%Using the techniques from \citet{Anisotropic} and \citet{yang2019spiked}, we find the asymptotic limit of $R(z)$ for any $z$ as $p$ goes to infinity, denoted by $R_\infty(z)$, with an almost optimal convergence rate of $p$.
%In particular, when $z=0$, the asymptotic limit of equation \eqref{lem_cov_shift_eq} is given by
%	\[ \tr[\Sigma_2^{-1/2} \Sigma \Sigma_2^{-1/2}R_\infty(0)]. \]
%
%For the bias asymptotic limit, we show in \todo{where?} that
%$$\bignorm{\Sigma_2^{1/2} ((X^{(1)})^{\top}(X^{(1)}) + (X^{(2)})^{\top}(X^{(2)})^{-1} w}^2= w^\top \Sigma_2^{-1/2}R'(0)\Sigma_2^{-1/2} w.$$
%Hence its limit can be calculated through $R_\infty'(z)$, which gives the expression in \eqref{lem_cov_derv_eq}.
%We leave the full proof of Lemma \ref{lem_cov_shift} and Lemma \ref{lem_cov_derivative} to Appendix \ref{sec_maintools}.
%Combining the above two results, we provide the proof of Theorem \ref{thm_main_informal} in Section \ref{app_proof_main_thm}.

>>>>

\medskip
\noindent{\bf Deriving the variance asymptotics.} Using definition \eqref{mainG}, we can write equation \eqref{eigen2extra} as
\be\label{rewrite X as R} [(X^{(1)})^\top X^{(1)}+(X^{(2)})^\top X^{(2)}]^{-1}=n^{-1}\Sigma_2^{-1/2}V\cal G(0)V^\top\Sigma_2^{-1/2}.\ee
When $z=0$, it is easy to check that equation \eqref{selfomega_a} can be reduced to \eqref{eq_a12extra}, which means that we actually have $a_1(0)=a_1$ and $a_2(0)=a_2$. Hence the matrix limit of $\cal G(0)$ is given by $(a_{1}\Lambda^2 + a_{2}\id_p)^{-1}$. Then inserting this limit into \eqref{rewrite X as R}, we can write the left-hand side of \eqref{lem_cov_shift_eq} as
\begin{align}
&\bigtr{\left( (X^{(1)})^{\top}X^{(1)} + (X^{(2)})^{\top}X^{(2)}\right)^{-1} \Sigma}\approx n^{-1}\bigtr{\Sigma_2^{-1/2}V\cal (a_{1}\Lambda^2 + a_{2}\id_p)^{-1}V^\top\Sigma_2^{-1/2}\Sigma}  \nonumber\\
&=n^{-1}\bigtr{\Sigma_2^{-1/2}\cal (a_{1}\Sigma_2^{-1/2}\Sigma_1\Sigma_2^{-1/2} + a_{2}\id_p)^{-1}\Sigma_2^{-1/2}\Sigma}  = n^{-1}\bigtr{\cal (a_{1} \Sigma_1  + a_{2}\Sigma_2)^{-1}\Sigma}  ,\label{Gi00}
\end{align}
where in the second step we used $V \Lambda^2 V^\top=M^\top M=\Sigma_2^{-1/2}\Sigma_1\Sigma_2^{-1/2}$ by equation \eqref{eigen2}.  This concludes \eqref{lem_cov_shift_eq} up to a small error.

@@

\medskip
\noindent{\bf Deriving the bias asymptotics.} Finally, we describe how to derive the bias asymptotics. We can write the left-hand side of equation \eqref{lem_cov_derv_eq} using the derivative of $\cal G$ with respect to $z$ at $z=0$. More precisely, using equation \eqref{eigen2extra} we can obtain that
\begin{align}
&n^2\bignorm{\Sigma_2^{1/2} \bigbrace{ (X^{(1)})^{\top}X^{(1)} + (X^{(2)})^{\top}X^{(2)} }^{-1} \Sigma_1^{1/2} w}^2 \nonumber\\
&=n^2 w^\top \Sigma_1^{1/2}\Sigma_2^{-1/2}V\left(   \Lambda U^\top (Z^{(1)})^\top Z^{(1)} U\Lambda  + V^\top (Z^{(2)})^\top Z^{(2)}V\right)^{-2}V^\top\Sigma_2^{-1/2}\Sigma_1^{1/2}w \nonumber\\
&=  w^\top \Sigma_1^{1/2}\Sigma_2^{-1/2}V\cal G'(0)V^\top\Sigma_2^{-1/2}\Sigma_1^{1/2}w,\label{calculate G'}
\end{align}
where we used equation \eqref{rewrite X as R} in the second step. Since the matrix limit of $\cal G(z)$ is given by \eqref{matrix limit}, it is natural to guess that the matrix limit of $\cal G'(0)$ is given by
\be\label{cal G'0}\cal G'(0) \approx \left.\frac{\dd}{\dd z}\right|_{z=0}(-z\id_p+a_{1}(z)\Lambda^2 + a_{2}(z)\id_p)^{-1} = \frac{\id_p- a_1'(0)\Lambda^2 - a_2'(0)\id_p}{(a_{1}(0)\Lambda^2 + a_{2}(0)\id_p)^2}.\ee
If we let $a_3:=-a_1'(0)$ and $a_4:=-a_2'(0)$, then taking implicit differentiation of equation \eqref{selfomega_a} we can check that $(a_3,a_4)$ satisfies equation \eqref{eq_a34extra}. Then inserting \eqref{cal G'0} into \eqref{calculate G'}, we obtain that
\begin{align}
& n^2\bignorm{\Sigma_2^{1/2} \bigbrace{ (X^{(1)})^{\top}(X^{(1)}) + (X^{(2)})^{\top}(X^{(2)}) }^{-1} \Sigma_1^{1/2} w}^2 \nonumber\\
&\approx  w^\top \Sigma_1^{1/2}\Sigma_2^{-1/2}V\frac{a_3\Lambda^2 +(1+ a_4)\id_p}{(a_{1}\Lambda^2 + a_{2}\id_p)^2}V^\top \Sigma_2^{-1/2}\Sigma_1^{1/2}w= w^{\top} \Pi_\bias w, \label{calculatePibias}
\end{align}
where in the last step we used $M = \Sigma_1^{1/2}\Sigma_2^{-1/2}$ and $V \Lambda^2 V^\top=M^\top M$. This concludes \eqref{lem_cov_derv_eq}. Note that in order to have the approximate identity for $\cal G'(0)$ in \eqref{cal G'0}, we not only need to know the asymptotics of $\cal G(0)$, but also need to know the asymptotics of $\cal G(z)$ for general $z$ around $z=0$. This is the main reason why we need to take a general $z$ in the definition of resolvents.
%In the above definition, we have taken the argument of $\cal G$ to be a general complex number, because we will need to use $\cal G'(0)$ in the proof of Lemma \ref{lem_cov_derivative}, which requires a good estimate of $\cal G(z)$ for $z$ around the origin.
%For the variance asymptotic limit, we study the resolvent
%	\[ R(z):= \bigbrace{\Sigma_2^{-1/2}( (X^{(1)})^{\top}(X^{(1)}) + (X^{(2)})^{\top}(X^{(2)}))\Sigma_2^{-1/2} - z \id}^{-1}, \text{ for any } z\in \C \text{ around } z=0. \]
%Using the techniques from \citet{Anisotropic} and \citet{yang2019spiked}, we find the asymptotic limit of $R(z)$ for any $z$ as $p$ goes to infinity, denoted by $R_\infty(z)$, with an almost optimal convergence rate of $p$.
%In particular, when $z=0$, the asymptotic limit of equation \eqref{lem_cov_shift_eq} is given by
%	\[ \tr[\Sigma_2^{-1/2} \Sigma \Sigma_2^{-1/2}R_\infty(0)]. \]
%
%For the bias asymptotic limit, we show in \todo{where?} that
%$$\bignorm{\Sigma_2^{1/2} ((X^{(1)})^{\top}(X^{(1)}) + (X^{(2)})^{\top}(X^{(2)})^{-1} w}^2= w^\top \Sigma_2^{-1/2}R'(0)\Sigma_2^{-1/2} w.$$
%Hence its limit can be calculated through $R_\infty'(z)$, which gives the expression in \eqref{lem_cov_derv_eq}.
%We leave the full proof of Lemma \ref{lem_cov_shift} and Lemma \ref{lem_cov_derivative} to Appendix \ref{sec_maintools}.
%Combining the above two results, we provide the proof of Theorem \ref{thm_main_informal} in Section \ref{app_proof_main_thm}.
